LLM-Driven CIDOC-CRM Knowledge Graph Extraction from Museum Archives

### 2025-05-04 paper/paper.tex
paper: Adze graphs Qwen and R1

%make sure in A4 paper
\usepackage[left=1.57cm,right=1.57cm,top=0.95cm,bottom=2.54cm]{geometry}

@@

@@

The results clearly illustrate a positive correlation between the parameter size of the language model and the resulting ontology conformance. Smaller models, such as Qwen 2.5 3B (48.63\%) and Llama 3 3B (52.28\%), demonstrated considerably low conformance rates. Increasing the parameter count led to improvements, as seen with Qwen 2.5 7B (50.79\%) and Llama 3 8B (63.97\%).

However, a significant performance gap exists between these smaller models and the larger DeepSeek models. DeepSeek V3 achieved a high score of 94.6\%, and DeepSeek R1 reached an impressive \textbf{98.73\%} ontology conformance. This suggests that accurately interpreting and applying ontology constraints remains a considerably more challenging task for smaller language models.

>>>>

%make sure in A4 paper
\usepackage[left=1.57cm,right=1.57cm,top=0.95cm,bottom=2.54cm]{geometry}
\usepackage{graphicx}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=3in]{assets/graph-qwen3b-adze}
    \caption{Ontology-guided graph from text "Adze" - Qwen2.5 3B}
    \label{fig_graph_qwen3b_adze}
\end{figure}

\begin{figure}[!t]
    \centering
    \includegraphics[width=3in]{assets/graph-r1-adze}
    \caption{Ontology-guided graph from text "Adze" - DeepSeek R1}
    \label{fig_graph_r1_adze}
\end{figure}

@@

The results clearly illustrate a positive correlation between the parameter size of the language model and the resulting ontology conformance. Smaller models, such as Qwen 2.5 3B (48.63\%) and Llama 3 3B (52.28\%) performed with considerably low conformance rates, while increasing the parameter count led to improvements as seen with Qwen 2.5 7B (50.79\%) and Llama 3 8B (63.97\%). Figures \ref{fig_graph_qwen3b_adze} and \ref{fig_graph_r1_adze} visually demonstrate this difference, showing example triplets extracted from the same source text about "Adze"\footnote{https://www.museumnasional.or.id/604/}.

However, a significant performance gap exists between these smaller models and the larger DeepSeek models. DeepSeek V3 achieved a high score of 94.6\%, and DeepSeek R1 reached an impressive \textbf{98.73\%} ontology conformance. This suggests that accurately interpreting and applying ontology constraints remains a considerably more challenging task for smaller language models compared to their larger counterparts.
